---
title: "R Notebook"
Author: Jia Wu
Date: 04/01/2020
output: html_notebook
description: analyze a school survey completed by parents and students in New York City
---
1. load libraries to read the data

school combined downloaded here:
https://data.world/dataquest/nyc-schools-data

Survey data downloaded here: https://data.cityofnewyork.us/Education/2011-NYC-School-Survey/mnz3-dyi8

```{r}
library(readr)
combined <- read_csv('2011 data files online/combined.csv')
general <- read_delim('2011 data files online/masterfile11_gened_final.txt', "\t")
d75 <- read_delim('2011 data files online/masterfile11_d75_final.txt', "\t")

```
2. Combine general and d74.
Select only the relevant variables: dbn, xx_xx_11 etc.
Rename column dbn to DBN.
```{r}
library(dplyr)
survey <- bind_rows(general, d75)
survey_2 <- survey %>%
  select(c(dbn,saf_p_11:aca_tot_11)) %>%
  rename(DBN=dbn)
```

3. Are there duplicates in survey_2
```{r}
duplicates <- survey_2 %>%
  mutate(is_dup = duplicated(DBN)) %>%
  filter(is_dup == TRUE)
```
4. who has the NA values:
most missing records were from student results
```{r}
colSums(is.na(survey_2))
```
5. What columns had zeros 
```{r}
colSums(survey_2==0)
```
6. only total counts has zeros. Should be change to NA.
Then confirm that there is no more 0 values in tot_xx_11 columns

```{r}
#install.packages("tidyr")
library(tidyr)
library(purrr)
library(dplyr)
#colSums(survey_2==0)
convert_to_NA <- function(vname) {
survey_2 <- survey_2 %>%
    mutate(!!vname := na_if(!!vname, 0))
    return(survey_2)
}
vname_list <- c(sym('saf_tot_11'), sym('com_tot_11'), sym('eng_tot_11'), sym('aca_tot_11'))
for (vname in vname_list) {
  survey_2 <- convert_to_NA(vname)
}
colSums(survey_2==0)
```
7. Reshape the survey database, rows = school x group (parent, student, teacher, total) x category ("saf", "com", "eng", "aca"), columns = ratings.

```{r}
library(ggplot2)
library(stringr)
category_names <- c('saf','com','eng','aca')
group_names <- c('_p_11', '_s_11', '_t_11', '_tot_11')

make_subframe <- function(category_name,group_names) {
  old_names <- map2_chr(category_name, group_names,str_c)
  new_names <- map2_chr("group", group_names,str_c)
  survey_temp <- survey_2 %>%
  select(DBN, !!old_names) %>%
  set_names(~str_to_lower(.) %>%
              str_replace_all(category_name,"group")) %>%
  pivot_longer(cols = !!new_names, names_to = "group", values_to = category_name)
  return(survey_temp)
}
survey_2_longer <- make_subframe(category_names[1],group_names)[,1:2]
for (category_name in category_names) {
  survey_2_longer <-cbind(survey_2_longer, make_subframe(category_name, group_names)[,3])
}

survey_2_longer2 <- survey_2_longer %>%
  pivot_longer(cols = !!category_names, names_to = "category", values_to = "rating")
```
8. Summarize parent, student, teacher ratings on each metrics

```{r}
survey_2_longer2 %>%
  group_by(category, group) %>%
  drop_na(rating)%>%
  summarize(mean(rating), sd(rating))
```
10. make boxplot for each metrics

```{r}
ggplot(data = survey_2_longer2)+
  aes(x = group, y = rating, color = group) + 
  geom_boxplot() + 
  facet_wrap(~category)
```

11. calculate the correlation of parent, student, teacher ratings
from observation, rating across different categories are small, and rating across different groups (parent, teacher, student) are large.
Should be able to confirm using repeated measures ANOVA
```{r}
cor_mat <- survey_2 %>%
select_if(is.numeric) %>%
cor(use = "pairwise.complete.obs")

cor_tib <- cor_mat %>%
  as_tibble(rownames = "variable")

```
12. sum categories together, leave groups separate.
Teacher correlates lower with parents and students, then between parents and students.
```{r}
survey_3 <- survey_2 %>%
mutate(parent_rating = saf_p_11+ com_p_11+ eng_p_11+ aca_p_11, student_rating = saf_s_11+ com_s_11+ eng_s_11+ aca_s_11, teacher_rating = saf_t_11+ com_t_11+ eng_t_11+ aca_t_11) %>%
  select(DBN,parent_rating, student_rating, teacher_rating)

cor_mat <- survey_3 %>%
select_if(is.numeric) %>%
cor(use = "pairwise.complete.obs")

cor_tib <- cor_mat %>%
  as_tibble(rownames = "group")

```
13. inner join survey_3 with combined to study the demographic influence of the surveys
```{r}
combined_survey <- survey_3 %>%
  inner_join(combined, by = "DBN")
```
14. study the correlations
parent_rating vs. Num of SAT Test Takers: r = -0.39
student_rating vs. high_score_percent: r = 0.46
parent and student_rating vs. average_class_size: r = -0.29, -0.26
parent_rating vs. total_enrollment: r = -0.33
student_rating vs. asian_per: r = 0.20
student_rating vs. black_per: r = -0.29
parent_rating vs. Total Cohort: r = -0.38
avg_sat_score is only slightly correlated with student rating r = 0.23, and not strongly correlated with parent or teacher rating.

```{r}
cor_mat <- combined_survey %>%
select_if(is.numeric) %>%
cor(use = "pairwise.complete.obs")

cor_tib <- cor_mat %>%
  as_tibble(rownames = "variable")

```

15. ratings in different boros
```{r}
combined_survey %>%
  select(parent_rating, student_rating, teacher_rating, boro) %>%
  drop_na()%>%
  group_by(boro) %>%
  summarize(mean(parent_rating), mean(student_rating), mean(teacher_rating))
```

16. boxplot comparing boros for different ratings
Parent from Staten Islan had the lowest rating among all the boros
```{r}
combined_survey2 <- combined_survey %>%
  select(parent_rating, student_rating, teacher_rating, boro) %>%
  drop_na(boro)%>%
  pivot_longer(cols = c(parent_rating, student_rating, teacher_rating), names_to = "group", values_to = "rating")

ggplot(data = combined_survey2) + 
  aes(x = boro, y = rating, color = boro) + 
  geom_boxplot()+
  facet_wrap(~group)
```

